\documentclass[11pt]{article}
\usepackage{graphicx}
\usepackage[utf8]{inputenc}
\usepackage{geometry}
\usepackage{titlesec}
\usepackage{enumitem}
\usepackage{hyperref} 
\usepackage{amsmath}
\usepackage{ gensymb }
\usepackage{ amssymb }
\usepackage{float}
\usepackage{amsthm}
\usepackage{float}

\newcommand{\R}{\mathbb{R}}
\newcommand{\N}{\mathbb{N}}
\newcommand{\Q}{\mathbb{Q}}
\newcommand{\Z}{\mathbff{Z}}
\newcommand{\st}{\text{s.t}}

\title{PSET5}
\author{Anthony Yoon}
\date{2/20/2025}
\begin{document}
\maketitle
\section{}
\subsection*{a}
We can first begin calulating MRS for each individual. Without a loss of generality, we can begin to note the following:
\[
MRS_i = \frac{\alpha (x_1^i)^{\alpha -1}(x_2^i)^\beta}{\beta(x_2^i)^{\beta - 1}(x_1^i)^\alpha}
\]
This above quantity implies that $\frac{x_2^1}{x_1^1} = \frac{x_2^2}{x_1^2}$. Let $e_1 = x_1^1 + x_1^2$ and $e_2 = x_2^1 + x_2^2$. Using this equalities as well as the implication derived from the MRSes, we can find that we get:
\[
x_2^1 = \frac{e_2}{e_1} x_1^1 \quad x_2^2 = \frac{e_2}{e_1} x_1^2
\]
Since both the above quantities are linear in nature with no intercept, this implies that indeed the contact curve is that of connecting endpoints. 
\subsection*{b}
Since we we are working with different utility functions, we can find that after similar caluclations to above that:
\[
MRS_1 = \frac{\alpha (x_1^1)^{\alpha - 1}(x_2^1)^{1-\alpha}}{(1-\alpha)(x_1^1)^\alpha (x_2^1)^{-\alpha}} = \frac{\alpha x_2^1}{(1-\alpha)x_1^1}
\]
and using similar calculations, we find that:
\[
MRS_2 = \frac{\beta x^2_2}{(1-\beta)x_1^2}
\]
Since we know that $1 > \alpha > \beta > 0$, we find that:
\[
\frac{\alpha}{1-\alpha} > \frac{\beta}{1-\beta}
\]
Thus, we can see that for MRS to equal to each other, we know that:
\[
\frac{x_2^1}{x_1^1} < \frac{x^2_2}{x^2_1}
\]
Using the equations derived above, we can find that:
\[
x_2^1 < \frac{e_2}{e_1} x_1^1
\]
this implies that the graph still intersects the origins, but now $x_2^1 < x_1^1$, where we have all a curve that will be strictly below that of the original line derived in \textbf{a}. The contract curve is seen as below. 
\subsection*{c}
For the contract curve to exist, we want $MRS_1 = MRS_2$. Let $e_1 = x_1^1 + x_1^2$. We can see that
\[
MRS_1 = MRS_2 \implies \alpha (x_1^1)^{\alpha - 1} = \beta (x_1^2)^{\beta - 1}
\]
Thus, substuting the endowment, we find that:
\[
    \alpha(x_1^1)^{\alpha - 1} = \beta (e_1 - x_1^1)^{\beta - 1}
\]
So we see that as $ x_1^1 \to e^1_1$, we find that the consumers will not consume any $x_2$, and consume only $x_1$ However, since we know that $\alpha > \beta$, this implies that consumer $x_1^1$ has greater value on $x_1^1$, which implies that $x_1^1  > x_1^2$. Note that any level of $x_2$ sastifies the MRS equality arguement, thus we are only concerned about when $e_1 = x_1^1 + x_1^2$ and one of these inputs are the utility maximizing solution. The contract curve will like the one below:
\section{}
\subsection*{a}
See Graph 
\subsection*{b}
See Graph 
\subsection*{c}
See Graph
\subsection*{d}
The core would just be the point $(e_1^1, e_2^1)$ as any trade would make this individual worse off, hence no trade would not be blocked by any coalition. 
\subsection*{e}
We first derive the general Marshallian Demand function for each individual. We are interested in the following optimization problem:
\begin{align*}
    \max & \quad x_1x_2\\
    \st & \quad p_1e_1 + p_2e_2 \geq p_1x_1 + p_2x_2
\end{align*}
Where the langrangian is:
\[
L = x_1x_2 - \lambda(p_1e_1 + p_2e_2 - p_1x_1 - p_2x_2)
\]
wherw we see that the FOCs are:
\begin{align*}
    [x_1] &\quad x_2 + \lambda p_1 \leq 0 \text{ and } x_1 \geq 0\\
    [x_2] & \quad x_1 + \lambda p_2 \leq 0 \text{ and } x_2 \geq 0\\
    [\lambda] &\quad p_1e_1 + p_2e_2 \leq p_1 x_1 + p_2x_2 
\end{align*}
We can see all FOCs must be strict equality, as if that is not the case, then markets will fail to clear $([x_1], [x_2])$ and by the AU assumption that we want to use all of our endowment to maximize utility. Using these FOCs, we can find that:
\[
p_1x_1 = p_2x_2
\]
Thus, using this equation and the constraint, we find that:
\[
x_1^m = \frac{e_1}{2} + \frac{p_2}{2p_1}e_2 \quad x_2^m = \frac{p_1}{2p_2}e_1 + \frac{e_2}{2}
\]
Thus, we find that:
\[
x_1^1 = \frac{p_2}{2p_1} + 1 \quad x_2^1 = \frac{p_1}{p_2} + \frac{1}{2}
\]
\[
x_1^2 = \frac{3p_2}{2p_1} + 1 \quad x_2^2 = \frac{p_1}{p_2} + \frac{3}{2}
\]
Thus we can find that where we let $\mathbf{p} = (p_1, p_2)$:
\[
z_1(\mathbf{p}) = 2 + 2 \left( \frac{p_2}{p_1} \right)  - 4 =  - 2 + 2 \left( \frac{p_2}{p_1} \right) 
\]
\[
z_2(\mathbf{p}) = 2 + 2 \left( \frac{p_1}{p_2} \right)  - 4 =  - 2 + 2 \left( \frac{p_1}{p_2} \right) 
\]
and we can verify that:
\[
\mathbf{p} \cdot \mathbf{z} = \begin{bmatrix}
    p_1 & p_2 
\end{bmatrix} \cdot 
\begin{bmatrix}
    2 \left( \frac{p_2}{p_1} \right) - 2\\ 
    2 \left( \frac{p_1}{p_2} \right) - 2
\end{bmatrix} = 2p_2 - 2p_1 + 2p_1 - 2p_2 = 0
\]
\subsection*{f}
We can see if $p_2 = p_1$, we find that obviously, $\mathbf{z}$ goes to 0. Thus, we find that the set of Walrasian equlibria is $p^*(\mathcal{E}) = \{p_1, p_2) | p_1 = p_2\}$ We find that a Walrasian equilbrium allocation is
\[
x^W =\{ (x_1, x_2)\} = \{ (1.5, 1.5), (2.5, 2.5)\}
\]
since there is one relative price, we can find that the set of Walrasian Equilbirum Allocations is: 
\[
W(\mathcal{E}) = \bigcup_{p^*} x^W(p^*(\mathcal{E}), \mathcal{E})  = \{ (1.5, 1.5), (2.5, 2.5)\}
\]
\subsection*{g}
Note that $e_1 = e_2$, this implies that the slope of the line is 1, which means that all allocations that have $x_1 = x_2$ will be in the core. Thus, we can see that the above set is a subset of the core. 
\subsection*{h}
Switching the utility function, we find that we are interested in the following optimization problem:
\begin{align*}
    \max & \quad x_1^{\frac{2}{3}}x_2^{\frac{1}{3}}\\
    \st & \quad p_1e_1 + p_2e_2 \leq p_1 x_1 + p_2x_2 
\end{align*}
the constraint remains the same, but with the following FOCs. 
\begin{align*}
    [x_1] & \quad \left( \frac{2}{3} x_1^{-\frac{1}{3}} \right)x_2^\frac{1}{3} = p_1 \lambda\\
    [x_2] & \quad \left( \frac{1}{3}x_1^{-\frac{1}{3}} \right)x_1^\frac{2}{3} = p_2 \lambda
\end{align*}
We know that these FOCs must have strict equality due to the same reasons as stated above. Using these FOCs, the following can be derived:
\[
2x_2p_2 = x_1p_1
\]
which implies that:
\[
x_1^2 = \frac{2}{3p_1} (p_1e_1 + p_2e_2) \quad x_2^2 = \frac{1}{3p_1}(p_1e_1 + p_2e_2)
\]
Using, previous results, we can find that:
\[
z_1(\mathbf{p}) = 1 + \frac{p_2}{2p_1} + \frac{4}{3} + \frac{2p_2}{p_1} - 4 = \frac{5p_2}{2p_1} - \frac{5}{3}
\]
\[
z_2(\mathbf{p}) = \frac{1}{2} + \frac{p_1}{p_2} + \frac{2p_1}{3p_2} + 1 -4 = \frac{5p_1}{3p_2} - \frac{5}{2}
\]
Note that this implies that $3p_2 = 2 p_1$, as this is the only relative price that makes Walras' law hold. Thus, we see that:
\[
x^W = \{(x_1, x_2)\} = \left\{ \left( \frac{4}{3}, 2 \right), \left( \frac{8}{3}, 2 \right)\right\}
\]
and since we have only one relative price, we can find that the set of Walrasian equilbirum is 
\[
    W(\mathcal{E}) = \bigcup_{p^*} x^W(p^*(\mathcal{E}), \mathcal{E})  =  \left\{ \left( \frac{4}{3}, 2 \right), \left( \frac{8}{3}, 2 \right)\right\}
\]
\subsection*{i}
Note that the optimal values of $x_1$ and $x_2$ are only dependent on prices, we are interested in solving the cases purely dependent on making the value of the endowments equal to each other. 
\subsubsection*{First Economy}
since we know that $p_1 = p_2$, and we are interested in equating market values to each other, we should know that:
\[
p_1(e_1^1 + e_2^1) = p_2(e_1^2 + p_2^2) \implies (e_1^1 + e_2^1) = (e_1^2 + p_2^2)
\]
Let $T = (T_1, T_2)$ where each $T_i$ represents the transfer from each individual 1 to individual 2 in respect to each good, we can see that:
\[
2 - T_1 + 1 - T_2 = 2 + T_1 + 3 + T_2 
\]
However, since prices are equal, we can solve with respect with each good. Thus, we can see that:
\[
2 - T_1 = 2 + T_1 \iff T_1 = 0 \quad 1 - T_2 = 3 + T_2 \iff T_2 = -1
\]
Thus, $T = (0, -1)$, which means that no transfer of good is done, but one $x_1$ is transfered from person 2 to person 1, with leads to:
\[
\mathbf{e^1} = \mathbf{e^2}= (2, 2)
\]
\subsubsection*{Second Economy}
Now. we see that $3p_2 = 2p_1$, using a similar logic to that above and same definition of $T$, we see that:
\begin{align*}
    p_1(2-T_1) + p_2(1-T_2) &= p_1(2+T_1) + p_2(3 + T_1)\\
    \frac{3p_2}{2}(2 - T_1) + p_2(1 - T_2) &= \frac{3p_2}{2}(2+T_1) + p_2(3+T_1)\\
\end{align*}
Using a similar arguement, we find that $T = (0, -1)$, which means that we obtain the same individual endowment as before.

\section{}
\subsection*{a}
See Graph
\subsection*{b}
Note that equating the MRSes here is not ideal due to the given utility functions, so a more logical arguement must be used. We can see that in order for each consumer to maximize their utility, consumer 1 and 2 should have only have good 1 and 2 respectively. Thus, the only pareto optimal points would be the bottom right corners of the edgeworth box, where we would draw the indfference curves as seen as below, making sure that consumer's 1 indifference curve goes through the bottom right corner of the indifference point and consumer's 2 indifference curve goes thorugh the top left corner of the Edgeworth Box. A graph can be seen below. 
\subsection*{c}
The core would be characterized as $\{e^1, e^2\} = \{(e_1^1, 0), (0, e_2^2) | e_1^1, e_2^2 \in [0,2]\}$. As moving to any allocation in this set will strictly increase utility for each consumer. 
\subsection*{d}
Consider 
\section{}
\subsection*{a}
We can note that as we are all using the minimum function, we can see that $x_1^1 = x_2^1 = x_2^2 = x_3^2 = x_1^3 = x_3^3$, which implies that we are interested in the allocation:
\[
    x_1^1 = x_2^1 = x_2^2 = x_3^2 = x_1^3 = x_3^3 = 0.5
\]
which is optimal as moving away would cause utilty lose and maximizes utlity for all and markets clear. Thus, this allocation is in the core, and let this allocation equal $\overline{x}$
\subsection*{b}
Consider the UMP for consumer 1. We are interested in the following optimization problem:
\begin{align*}
    \max & \quad \min\{x_1^1, x_2^1\}\\
    \st & \quad x_1^1 p_1 + p_1x_2^1 \leq p e_1
\end{align*}
note that we want $x_1^1 = x_2^1$, which implies using the strict equality of the constraint (which we know is true by the AU assumption) that 
\[
x_1^1 = x_2^1 = \frac{p_1}{p_1 + p_2}
\]
using a similar logic, we can see that:
\[
x_2^2 = x_3^2 = \frac{p_2}{p_2+p_3}
\]
and
\[
x_1^3 = x_3^3 = \frac{p_3}{p_1+p_3}
\]
Thus, we can now analyze the aggegate demand functions. 
We find that $z_i = x_i^1 + x_i^2 + x_i^3 - 1$ where $i \in \{1,2,3\}$. After some algebra and using the Marshallian demand functions taht we derived, we find that $p_1 = p_2 = p_3$. This implies that:
\[
    x_1^1 = x_2^1 = x_2^2 = x_3^2 = x_1^3 = x_3^3 = 0.5
\]
Thus, we can see that $p^*(\mathcal{E}) = \{(p_1, p_2, p_3) \in \R^3 | p_1 = p_2 = p_3\}$. Thus, we can see tha $x^W(p^*(\mathcal{E}), \mathcal{E}) = \{ (0.5,0.5,0), (0, 0.5, 0.5), (0.5, 0, 0.5)\}$ Since we are working with only one relative price, we can note that 
\[
    W(\mathcal{E}) = \bigcup_{p^*} x^W(p^*(\mathcal{E}), \mathcal{E}) = \{ (0.5,0.5,0), (0, 0.5, 0.5), (0.5, 0, 0.5)\}
\]
\section*{6}

\end{document}
